Galactic Chemical Evolution Simulations Favor Long Delay Times for Type Ia Supernovae

### 2024-01-16 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class
\documentclass[modern,linenumbers]{aastex631}
% \documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
%twocolappendix

@@

% [Hook and intro paragraph]
Galactic chemical evolution studies seek to reproduce the observed distribution of metals as a function of age throughout the Milky Way Galaxy. \citet{Tinsley1979-StellarLifetimes} proposed that the non-Solar \aFe ratios seen by, e.g., \citet{Wallerstein1962-GDwarfAbundances} were caused by different stellar lifetimes for the contributors of the Fe-peak elements than the $\alpha$-elements. Type Ia supernovae (SNe Ia), the explosions of carbon-oxygen white dwarfs (WDs), are responsible for roughly half of the iron produced in the Galaxy [cite]. SNe Ia have long been known to occur in elliptical galaxies, suggesting that their progenitors may have lifetimes of up to many Gyr. \citet{Tinsley1979-StellarLifetimes} assumed a range of $\sim 3\times 10^7 - 10^8$ years for the lifetimes of SN Ia progenitors, although they argued against the hypothesis that SNe Ia are exploding white dwarfs. \citet{Matteucci1986-SupernovaEnrichment} modeled SNe Ia as the end product of binary evolution and arrived at much longer lifetimes for the progenitors, finding that SNe Ia did not contribute substantially to chemical enrichment for $\sim 10^8$ years after the commencement of star formation. 
% [Concluding sentence about how understanding the DTD is important for chemical evolution and for understanding the progenitors of SNe Ia]
The timescale for Fe enrichment in the Galaxy, a key component of chemical evolution models, is directly linked to the progenitor lifetimes of SNe Ia.

% [What is the delay time distribution]
The delay-time distribution (DTD) is the rate of SNe Ia which explode as a function of time after a hypothetical instantaneous star formation event. By convolving the DTD with the star formation rate (SFR) as a function of time, one can produce a function describing the time-variation in the SN Ia rate.
Different models for the SN Ia progenitor have different DTDs.

% [Introduction to SNe Ia]
% Type Ia supernovae (SNe Ia), the explosions of carbon-oxygen white dwarfs (WDs), are responsible for roughly half of the iron produced in the Galaxy [cite]. Despite their importance, 
The mechanism(s) for producing SNe Ia are not fully understood \citep[for reviews, see][]{Maoz2014-Review,Livio2018-ProgenitorReview,Ruiter2020-ProgenitorReview}. Two general production channels have been proposed: in the single-degenerate (SD) case, the WD accretes mass from a close non-degenerate companion until it surpasses a mass of $\sim1.4$ M$_\odot$ and explodes \citep{Whelan1973-SDModel,Nomoto1982-SDModel,Yoon2003-SDModel}. In the double-degenerate (DD) case, two WDs merge after a gravitational-wave inspiral \citep{Iben1984-IaBinary,Webbink1984-DDModel,Pakmor2012-WDMerger} or head-on collision \citep{Benz1989-CollisionalDD,Thompson2011-CollisionalDD}. Recent observations have placed tight constraints on the SD channel, heavily disfavoring it as the main pathway for producing ``normal'' SNe Ia \citep[e.g.,][]{Panagia2006-RadioEmission,Schaefer2012-ExCompanionSNR,Chomiuk2016-RadioEmission,Fausnaugh2019-EarlyIaLightCurves,Tucker2020-SNeIaSpectra,Do2021-Progenitor1972E,Tucker2023-SN2011fe}. Though it is now the preferred model, the DD channel faces issues of its own with matching observed SN Ia rates because not all WD mergers may lead to a thermonuclear explosion \citep[e.g.,][]{Shen2012-DDMergers}.

[Analytical DTDs]
Uncertainties in the progenitor channels of SNe Ia translate into uncertainties in the form of the DTD. 
\citet{Greggio2005-AnalyticalRates} used analytical approximations of binary population synthesis to produce theoretical DTDs for the SD and DD channels.

@@

[Why radial migration is important for SNe Ia]
Core-collapse SNe explode on timescales of a few Myr, before the progenitor has a chance to migrate far from its birth location. On the other hand, SNe Ia have a broad distribution of delay times and their progenitors can migrate great distances between birth and explosion. The consequence is that SNe Ia can enrich regions of the Galaxy far removed from regions which have experienced a high rate of star formation.

@@

In this paper we comprehensively evaluate a selection of DTD models from the literature with multiple SFHs and a prescription for radial migration in the Versatile Integrator for Chemical Evolution (VICE). In Section \ref{sec:methods}, we present our models for the DTD and SFH, and we describe our chemical evolution simulations. In Section \ref{sec:onezone-results}, we present the results of our one-zone chemical evolution simulations. In Section \ref{sec:multizone-results}, we present the results of our multi-zone simulations and compare to observations. In Section \ref{sec:discussion}, we discuss the implications for the DTD and future surveys. In Section \ref{sec:conclusions}, we conclude.

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit parameters which have been left unchanged from \citet{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone simulation parameters
        R_{\rm gal}     & [0, 20] kpc   & & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & & Maximum simulation time \\
        n               & 8         & & Number of stellar populations formed per ring per time-step \\
        R_{\rm Ia}(t)   & Equations \ref{eq:powerlaw-dtd}--\ref{eq:triple-dtd}  & \ref{sec:dtd-models}  & Delay time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        M_{\rm gas,0}   & 0         & & Initial gas mass \\
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & & SN Ia yield of Fe \\
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & & Time-dependence of the SFR in the inside-out model \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & & Time-dependence of the SFR in the late-burst model \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & & Time-dependence of the IFR in the early-burst model \\
        \uptau_{*,\rm EB}(t|R_{\rm gal})  & Equation \ref{eq:earlyburst-conroy22} & & Time-dependence of the SFE timescale in the early-burst model \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & & Time-dependence of the IFR in the two-infall model \\
    \enddata
\end{deluxetable*}

@@

\begin{subequations}
\begin{align}
    \dot M_{\rm Fe}^{\rm Ia} &= y_{\rm Fe}^{\rm Ia} \langle \dot M_*\rangle_{\rm Ia} \\
    &= y_{\rm Fe}^{\rm Ia} \frac{\int_0^t \dot M_*(t')R_{\rm Ia}(t-t')dt'}{\int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt'}
\end{align}
\end{subequations}

The yield $y_{\rm Fe}^{\rm Ia}$ can be expressed in terms of the integral of the DTD:

\begin{equation}
    y_{\rm Fe}^{\rm Ia} = m_{\rm Fe}^{\rm Ia} \int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt' = m_{\rm Fe}^{\rm Ia} \frac{N_{\rm Ia}}{M_*}
    \label{eq:dtd-integral}
\end{equation}
where $m_{\rm Fe}^{\rm Ia}$ is the average mass of Fe produced by a single SN Ia. We take the average number of SNe Ia per mass of stars formed to be $N_{\rm Ia}/M_*=2.2\times10^{-3}\,{\rm M}_\odot^{-1}$ following \citet{MaozMannucci2012-SNeIaReview}, which gives $R_{\rm Ia}$ units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$.

@@

\begin{figure}
    \centering
    \includegraphics{figures/delay_time_distributions.pdf}
    \caption{Selection of models for the SN Ia delay time distribution (DTD) used in this paper. All functions are normalized such that $R_{\rm Ia}(t=1\,{\rm Gyr})=1$. The black squares represent the DTD recovered for the SDSS-II sample of SNe Ia by \citet{Maoz2012-SloanIIDTD} at the same scale as the DTD models. The horizontal and vertical error bars indicate the time range and 1$\sigma$ uncertainties, respectively, of each DTD measurement. The colored circles along the horizontal axis indicate the median delay time for each model in the same order as the legend.}
    \label{fig:dtds}
    \script{delay_time_distributions.py}
\end{figure}

\begin{deluxetable*}{lll}
\tablecaption{Summary of SN Ia DTD models explored in this paper.\label{tab:dtds}}
\tablehead{
\colhead{Model} & \colhead{Parameters} & \colhead{Similar to}
}
\startdata
Power-law   & $\alpha=-1.1$                 & \citet[][field]{Maoz2017-CosmicDTD}; 
                                              \citet{Wiseman2021-DESRates}              \\
Power-law   & $\alpha=-1.4$                 & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                              \citet{Heringer2019-FieldGalaxyDTD}       \\
Plateau     & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
Plateau     & $W=1.0$ Gyr, $\alpha=-1.1$    & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
Triple-system   & $f_{\rm init}=0.05f_{\rm peak}$, $t_{\rm rise}=0.5$ Gyr, & \citet{Rajamuthukumar2023-TripleEvolution} \\
            & $W=0.5$ Gyr, $\alpha=-1.1$ & \\
Exponential & $\tau=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                  \citet{Schonrich2009-RadialMixing};       \\
            &                   & \citet{Weinberg2017-ChemicalEquilibrium}  \\
Exponential & $\tau=3.0$ Gyr    & --- \\
Two-population  & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, & \citet{Mannucci2006-TwoPopulations} \\
            & $\tau=3.0$ Gyr & \\
\enddata
\end{deluxetable*}

@@

In this paper, we consider four models for the SFH: inside-out, late-burst, early-burst, and two-infall. The first two models are run in VICE's ``star formation mode,'' where the surface density of star formation rate $\dot\Sigma_*$ is prescribed along with the star formation efficiency timescale $\tau_*$, and the infall rate surface density $\dot\Sigma_{\rm in}$ and gas surface density $\Sigma_{\rm gas}$ are calculated from the specified quantities. The latter two models are run in ``infall mode,'' where we specify $\dot\Sigma_{\rm in}$  and $\tau_*$ and the other two quantities follow.

@@

All models are run for a duration of $t_{\rm max}=13.2$ Gyr with a time-step of $\Delta t=0.01$ Gyr. We adopt an outflow mass-loading factor $\eta\equiv \dot M_{\rm out}/\dot M_*=2.15$ \citep[see Equation 8 from][]{Johnson2021-Migration}, a star formation efficiency timescale $\tau_*\equiv M_{\rm gas}/\dot M_*=2$ Gyr, and an initial gas mass $M_{\rm gas,0}=0$. We adopt a continuous recycling prescription \citep[see Equation 2 from][]{JohnsonWeinberg2020-Starbursts}. Unless otherwise specified, we adopt the inside-out star formation rate (Equation \ref{eq:insideout-sfh}) evaluated at $R_{\rm gal}=8$ kpc (i.e., $\tau_{\rm rise}=2$ Gyr and $\tau_{\rm sfh}=15.1$ Gyr), and a minimum SN Ia delay time of 40 Myr. A summary of these one-zone fiducial parameters is presented in Table \ref{tab:onezone-parameters}.

\begin{deluxetable}{Ccl}
    \tablecaption{A summary of parameters and their fiducial values for our one-zone chemical evolution models.\label{tab:onezone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Value} & \colhead{Description}
    }
    \startdata
    \Delta t        & 10 Myr    & Time-step size \\
    t_{\rm max}     & 13.2 Gyr  & Maximum simulation time \\
    \eta            & 2.15      & Outflow mass-loading factor \\
    \tau_*          & 2.0 Gyr   & Star formation efficiency timescale \\
    \dot M_r        & continuous    & Recycling rate \\
    t_D             & 40 Myr    & Minimum SN Ia delay time \\
    M_{\rm gas,0}   & 0         & Initial gas mass \\
    \dot M_*(t)     & Equation \ref{eq:insideout-sfh}  & Star formation rate \\
    \tau_{\rm rise} & 2 Gyr     & SFR rise timescale \\
    \tau_{\rm sfh}  & 15.1 Gyr  & SFR exponential timescale
    \enddata
\end{deluxetable}

@@

@@

@@

>>>>

% Define document class
% \documentclass[modern,linenumbers]{aastex631}
\documentclass[twocolumn,twocolappendix,linenumbers,trackchanges]{aastex631}
%twocolappendix

@@

% [Hook and intro paragraph]
Galactic chemical evolution studies seek to explain the observed distribution of metals %as a function of age 
throughout the Milky Way Galaxy. \citet{Tinsley1979-StellarLifetimes} proposed that the non-Solar \aFe\footnote{
    In standard bracket notation, $[X/Y]\equiv \log_{10}(X/Y) - \log_{10}(X/Y)_{\odot}$. In this work we will use \aFe and [O/Fe] interchangeably, although observational studies will often use a combination of $\alpha$-elements to calculate \aFe.
} ratios seen by, e.g., \citet{Wallerstein1962-GDwarfAbundances} were caused by different stellar lifetimes for the contributors of the Fe-peak elements than the $\alpha$-elements. Type Ia supernovae (SNe Ia), the explosions of carbon-oxygen white dwarfs (WDs), are responsible for a majority of the Fe produced in the Galaxy \citep{Matteucci1986-SupernovaEnrichment}; meanwhile, core collapse supernovae (CCSNe), the explosions of massive stars, produce the $\alpha$-elements (e.g., O and Mg) in addition to a smaller fraction of Fe. SNe Ia have long been known to occur in elliptical galaxies, suggesting that their progenitors may have lifetimes of up to many Gyr. 
% \citet{Tinsley1979-StellarLifetimes} assumed a range of $\sim 3\times 10^7 - 10^8$ years for the lifetimes of SN Ia progenitors, although they argued against the hypothesis that SNe Ia are exploding white dwarfs. 
\citet{Matteucci1986-SupernovaEnrichment} modeled SNe Ia as the end product of binary evolution and %arrived at much longer lifetimes for the progenitors, finding 
concluded that SNe Ia did not contribute substantially to chemical enrichment for $\sim 10^8$ years after the commencement of star formation. The timescale for Fe enrichment in the Galaxy, a key component of chemical evolution models, is directly linked to the progenitor lifetimes of SNe Ia. 
Therefore, the balance of SN rates over time regulates the relative abundances of the $\alpha$-elements and Fe as a function of stellar age. [Better concluding sentence]

% [What is the delay time distribution]
SNe Ia are delayed by $\sim10^8 - 10^{10}$ years after the formation of their progenitors as a result of their stellar and binary evolution. The delay-time distribution (DTD) is the rate of SNe Ia which explode as a function of time after the formation of a hypothetical unit-mass single stellar population. When the DTD is convolved with the Galactic star formation rate (SFR) as a function of time, it yields the SN Ia rate as a function of time. The DTD controls the delay between chemical enrichment from CCSNe, which explode on timescales of a few Myr (smaller than the timestep size of many chemical evolution simulations), and SNe Ia. As such, it is a key parameter in the chemical evolution of the Galaxy. The DTD reflects the evolution of the SNe Ia progenitor systems, so different models for the progenitors of SNe Ia will naturally predict different forms for the DTD.

% [Introduction to SNe Ia]
The mechanism(s) for producing SNe Ia are not fully understood \citep[for reviews, see][]{Maoz2014-Review,Livio2018-ProgenitorReview,Ruiter2020-ProgenitorReview}. Two general production channels have been proposed: in the single-degenerate (SD) case, the WD accretes mass from a close non-degenerate companion until it surpasses a mass of $\sim1.4$ M$_\odot$ and explodes \citep{Whelan1973-SDModel,Nomoto1982-SDModel,Yoon2003-SDModel}. In the double-degenerate (DD) case, two WDs merge after a gravitational-wave inspiral \citep{Iben1984-IaBinary,Webbink1984-DDModel,Pakmor2012-WDMerger} or head-on collision \citep{Benz1989-CollisionalDD,Thompson2011-CollisionalDD}. Recent observations have placed tight constraints on the SD channel, heavily disfavoring it as the main pathway for producing ``normal'' SNe Ia \citep[e.g.,][]{Panagia2006-RadioEmission,Schaefer2012-ExCompanionSNR,Chomiuk2016-RadioEmission,Fausnaugh2019-EarlyIaLightCurves,Tucker2020-SNeIaSpectra,Do2021-Progenitor1972E,Tucker2023-SN2011fe}. Though it is now the preferred model, the DD channel faces issues of its own with matching observed SN Ia rates because not all WD mergers may lead to a thermonuclear explosion \citep[e.g.,][]{Shen2012-DDMergers}.

% [Analytical DTDs]
Uncertainties in the progenitor channels of SNe Ia translate into uncertainties in the form of the DTD. \citet{Greggio2005-AnalyticalRates} formulated analytical DTDs for several progenitor models; she found that the SD channel tends to produce a steeper DTD function than the DD channel, and the median delay time can vary by a factor of 10 depending on the chosen progenitor scenario and its particular parameters.

@@

% [Why radial migration is important for SNe Ia]
A complicating factor in observing and modeling the effects of the DTD is the migration of stars throughout the disk. \citet{SellwoodBinney2002-RadialMixing} first demonstrated that a transient spiral perturbation (or other potential feature; see [citations]) can cause resonant interactions which adjust stars' guiding center radii without leading to kinematic heating of the disk. Core-collapse SNe explode on timescales of a few Myr, before the progenitor has a chance to migrate far from its birth location. On the other hand, SNe Ia have a broad distribution of delay times and their progenitors can migrate great distances between birth and explosion. The consequence is that SNe Ia can enrich regions of the Galaxy far removed from regions which have experienced a high rate of star formation. The first GCE model to incorporate radial mixing was developed by \citet{Schonrich2009-RadialMixing}, and subsequent studies have incorporated more comprehensive numerical treatments of radial migration and the effect of a central bar \citep[e.g.,][]{Minchev2013-ChemodynamicalEvolution}. In this paper, we follow the treatment by \citet{Johnson2021-Migration}, who [describe treatment].

@@

In this paper we comprehensively evaluate a selection of DTD models from the literature with multiple SFHs and a prescription for radial migration in the Versatile Integrator for Chemical Evolution \citep[VICE;][]{JohnsonWeinberg2020-Starbursts}. In Section \ref{sec:methods}, we present our models for the DTD and SFH, and we describe our chemical evolution simulations. In Section \ref{sec:onezone-results}, we present the results of our one-zone chemical evolution simulations. In Section \ref{sec:multizone-results}, we present the results of our multi-zone simulations and compare to observations. In Section \ref{sec:discussion}, we discuss the implications for the DTD and future surveys. In Section \ref{sec:conclusions}, we conclude.

@@

We make use of the multi-zone simulation tools in VICE developed by \citet{Johnson2021-Migration}. The basic setup of our models follows theirs: the simulated Galaxy is a disk with a radius of 20 kpc and is divided into concentric rings of width $\delta R_{\rm gal}=100$ pc. The gas in each ring is assumed to be instantaneously mixed, and no gas is exchanged between rings. We run our simulations with a time-step size of $\Delta t=10$ Myr up to a maximum time of $t_{\rm max}=13.2$ Gyr. Following \citet{Johnson2021-Migration}, we set VICE to form $n=8$ stellar populations per ring per time-step, and we set a maximum star-formation radius of $R_{\rm SF} = 15.5$ kpc, such that $\dot\Sigma_*=0$ for $R_{\rm gal}>R_{\rm SF}$.

Table \ref{tab:multizone-parameters} summarizes many of the parameters in our multi-zone models.

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit parameters which have been left unchanged from \citet{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone simulation parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:methods} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:methods} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:methods} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:methods} & Maximum simulation time \\
        n               & 8         & \ref{sec:methods} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:methods} & Maximum radius of star formation \\
        M_{\rm gas,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        % DTD
        R_{\rm Ia}(t)   & Equations \ref{eq:powerlaw-dtd}--\ref{eq:prompt-dtd}  & \ref{sec:dtd-models}  & Delay time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_*  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the SFR in the inside-out model \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the SFR in the late-burst model \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the IFR in the early-burst model \\
        \uptau_{*,\rm EB}(t|R_{\rm gal})  & Equation \ref{eq:earlyburst-conroy22} & \ref{sec:sfh}   & Time-dependence of the SFE timescale in the early-burst model \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the IFR in the two-infall model \\
    \enddata
\end{deluxetable*}

\subsection{Nucleosynthetic Yields}
\label{sec:yields}

CCSN enrichment:
\begin{equation}
    \dot M_x^{\rm CC} = y_x^{\rm CC} \dot M_*
    \label{eq:ccsn-enrichment}
\end{equation}
where $y_x^{\rm CC}$ is the CCSN yield of element $x$ per unit mass of star formation. Following \citet{Johnson2021-Migration}, who in turn adopt their yields from \citet{Weinberg2017-ChemicalEquilibrium}, we adopt $y_{\rm O}^{\rm CC}=0.015$ and $y_{\rm Fe}^{\rm CC}=0.0012$.

We follow the formalism of \citet{Weinberg2017-ChemicalEquilibrium} when deriving the equations for Fe enrichment from SNe Ia.

$R_{\rm Ia}(t)$ is the DTD in units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$.

The rate of Fe contribution to the ISM from SNe Ia is 
\begin{equation}
    \dot M_{\rm Fe}^{\rm Ia} = y_{\rm Fe}^{\rm Ia} \langle \dot M_*\rangle_{\rm Ia} \\
\end{equation}
where $\langle \dot M_*\rangle_{\rm Ia}$ is the time-averaged SFR weighted by the DTD and $y_{\rm Fe}^{\rm Ia}$ is the Fe yield of SNe Ia. \citet{Weinberg2017-ChemicalEquilibrium} show in their Appendix A that
\begin{equation}
    \langle \dot M_*\rangle_{\rm Ia} \equiv \frac{\int_0^t \dot M_*(t')R_{\rm Ia}(t-t')dt'}{\int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt'}
    \label{eq:weighted-sfr}
\end{equation}
where $\dot M_*(t)$ is the SFR at time $t$, $R_{\rm Ia}(t) dt$ is the number of SNe Ia which explode in the time interval $(t, t+dt)$ per unit mass of stars formed at $t=0$, $t_D$ is the minimum SN Ia delay time, and $t_{\rm max}$ is the maximum simulation time. The denominator of Equation \ref{eq:weighted-sfr} is therefore the total number of SNe Ia per M$_{\odot}$ of stars formed.

The yield $y_{\rm Fe}^{\rm Ia}$ measures the mass of Fe produced by SNe Ia per unit mass of star formation. It can be expressed in terms of the integral of the DTD:
\begin{equation}
    y_{\rm Fe}^{\rm Ia} = m_{\rm Fe}^{\rm Ia} \int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt' = m_{\rm Fe}^{\rm Ia} \frac{N_{\rm Ia}}{M_*}
    \label{eq:dtd-integral}
\end{equation}
where $m_{\rm Fe}^{\rm Ia}$ is the average mass of Fe produced by a single SN Ia. We take the average number of SNe Ia per mass of stars formed to be $N_{\rm Ia}/M_*=2.2\times10^{-3}\,{\rm M}_\odot^{-1}$ following \citet{MaozMannucci2012-SNeIaReview}, which gives $R_{\rm Ia}$ units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$. Following \citet{Johnson2021-Migration}, we take $y_{\rm Fe}^{\rm Ia}=0.00214$. SNe Ia do not contribute O so $y_{\rm O}^{\rm Ia}=0$.

@@

@@

\begin{figure}
    \centering
    \includegraphics{figures/delay_time_distributions.pdf}
    \caption{Selection of models for the SN Ia delay time distribution (DTD) used in this paper. All functions are normalized such that $R_{\rm Ia}(t=1\,{\rm Gyr})=1$. The black squares represent the DTD recovered for the SDSS-II sample of SNe Ia by \citet{Maoz2012-SloanIIDTD} at the same scale as the DTD models. The horizontal and vertical error bars indicate the time range and 1$\sigma$ uncertainties, respectively, of each DTD measurement. The colored circles along the horizontal axis indicate the median delay time for each model.}
    \label{fig:dtds}
    \script{delay_time_distributions.py}
\end{figure}

\begin{deluxetable*}{lcll}
\tablecaption{Summary of SN Ia DTD models explored in this paper.\label{tab:dtds}}
\tablehead{
\colhead{Model} & \colhead{Eq.} & \colhead{Parameters} & \colhead{Similar to}
}
\startdata
Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.1$                 & \citet[][field]{Maoz2017-CosmicDTD}; 
                                              \citet{Wiseman2021-DESRates}              \\
Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.4$                 & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                              \citet{Heringer2019-FieldGalaxyDTD}       \\
Plateau     & \ref{eq:plateau-dtd}  & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
Plateau     & \ref{eq:plateau-dtd}  & $W=1.0$ Gyr, $\alpha=-1.1$    & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
Triple-system   & \ref{eq:triple-dtd}   & $f_{\rm init}=0.05f_{\rm peak}$, $t_{\rm rise}=0.5$ Gyr, & \citet{Rajamuthukumar2023-TripleEvolution} \\
                &                       & $W=0.5$ Gyr, $\alpha=-1.1$ & \\
Exponential & \ref{eq:exponential-dtd}  & $\tau=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                                              \citet{Schonrich2009-RadialMixing};       \\
            &                           &                   & \citet{Weinberg2017-ChemicalEquilibrium}  \\
Exponential & \ref{eq:exponential-dtd}   & $\tau=3.0$ Gyr    & --- \\
Two-population  & \ref{eq:prompt-dtd}   & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, & \citet{Mannucci2006-TwoPopulations} \\
                &                       & $\tau=3.0$ Gyr & \\
\enddata
\end{deluxetable*}

@@

In this paper, we consider four models for the SFH: inside-out, late-burst, early-burst, and two-infall. The first two models are run in VICE's ``star formation mode,'' where the surface density of star formation rate $\dot\Sigma_*$ is prescribed along with the star formation efficiency timescale $\tau_*$, and the infall rate surface density $\dot\Sigma_{\rm in}$ and gas surface density $\Sigma_{\rm gas}$ are calculated from the specified quantities. The latter two models are run in ``infall mode,'' where we specify $\dot\Sigma_{\rm in}$  and $\tau_*$ and the other two quantities follow. All models have an initial gas mass $M_{\rm gas,0} = M_{\rm gas}(t=0) = 0$ in all zones.

@@

% [Intro paragraph, motivation for running one-zone models]
Before running the full multi-zone simulations, it will be useful to observe the effects of the DTD in more idealized conditions. A one-zone model assumes the entire gas reservoir is instantaneously mixed, so there is no spacial dependence for the distribution of metal abundances. In this section

All models are run for a duration of $t_{\rm max}=13.2$ Gyr with a time-step of $\Delta t=0.01$ Gyr. We adopt the inside-out star formation rate (Equation \ref{eq:insideout-sfh}) evaluated at $R_{\rm gal}=8$ kpc (i.e., $\tau_{\rm rise}=2$ Gyr and $\tau_{\rm sfh}=15.1$ Gyr), a star formation efficiency timescale $\tau_*\equiv M_{\rm gas}/\dot M_*=2$ Gyr, and an initial gas mass $M_{\rm gas,0}=0$. We adopt a continuous recycling prescription \citep[see Equation 2 from][]{JohnsonWeinberg2020-Starbursts}. Unless otherwise specified, we adopt an outflow mass-loading factor $\eta\equiv \dot M_{\rm out}/\dot M_*=2.15$ \citep[see Equation 8 from][]{Johnson2021-Migration} and a minimum SN Ia delay time of 40 Myr. A summary of these one-zone fiducial parameters is presented in Table \ref{tab:onezone-parameters}.
\begin{deluxetable}{Ccl}
    \tablecaption{A summary of parameters and their fiducial values for our one-zone chemical evolution models.\label{tab:onezone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Value} & \colhead{Description}
    }
    \startdata
    \Delta t        & 10 Myr    & Time-step size \\
    t_{\rm max}     & 13.2 Gyr  & Maximum simulation time \\
    \eta            & 2.15      & Outflow mass-loading factor \\
    \tau_*          & 2.0 Gyr   & Star formation efficiency timescale \\
    \dot M_r        & continuous    & Recycling rate \\
    t_D             & 40 Myr    & Minimum SN Ia delay time \\
    M_{\rm gas,0}   & 0         & Initial gas mass \\
    \dot M_*(t)     & Equation \ref{eq:insideout-sfh}  & Star formation rate \\
    \tau_{\rm rise} & 2.0 Gyr     & SFR rise timescale \\
    \tau_{\rm sfh}  & 15.1 Gyr  & SFR exponential timescale
    \enddata
\end{deluxetable}

@@

[Subsection introduction]
Section \ref{sec:onezone-results} and specifically Figures \ref{fig:onezone-threepanel} and \ref{fig:onezone-twopanel} illustrated the effect of the DTD models and parameters on [O/Fe]--[Fe/H] tracks in idealized one-zone models.

@@

[Subsection introduction]

@@

Figure \ref{fig:age-ofe-dtd} similarly plots the simulated distributions in age--[O/Fe] space for five of our DTD models. All simulations were run with the early-burst SFH model because the high-$\alpha$ sequence was distinct in metallicity space in Figure \ref{fig:age-ofe-sfh}. Similar to Figure \ref{fig:ofe-feh-dtd}, simulations are arranged from left to right by increasing median SN Ia delay time. Moving from left to right, the high-$\alpha$ sequence moves to higher [O/Fe] with increasing median delay time, from $\sim0.4$ for the two-population model to $\sim0.4$ for the triple-system DTD. As we have seen in previous figures, the range in [O/Fe] produced by the models on the left is much smaller than for the models on the right. At high $|z|$ (top row), the observed range of [O/Fe] is larger than produced by any of our simulations, although the plateau ($W=1$ Gyr) and triple-system models come close. While the other three fall short of the observed range in [O/Fe], they more closely match the median age--[O/Fe] relation. There is a slight reversal in the observed trend for the stars with the highest [O/Fe]: the $0.45\leq{\rm [O/Fe]}<0.5$ bin has a slightly \textit{lower} median age than the $0.3\leq{\rm [O/Fe]}<0.35$ bin at high $|z|$ in the \citet{Leung2023-Ages} sample, a small effect but one which is not observed in any of our simulation results.
